---
title: "Reproducible Research: Peer Assessment 1"
author: "Karna Desai"
date: "December 20, 2015"
output: html_document
keep_md: true
---
## Load Data
=========

```{r, echo=TRUE}
library(lattice)
library(knitr)
library(ggplot2)

data <- read.csv("activity.csv")
```

## Mean total number of steps taken per day:

```{r, echo=TRUE}

steps_by_day <- aggregate(steps ~ date, data, sum)
hist(steps_by_day$steps, main = paste("Total Steps Each Day (excluding: missing values)"), col="blue",breaks = 10, xlab="Number of Steps")
Step_mn <- mean(steps_by_day$steps)
Step_md <- median(steps_by_day$steps)
Step_mn
Step_md
```

The mean and median total number of steps taken per day are 10766.19 and 10765 respectively. 


## What is the average daily activity pattern?

```{r, echo=TRUE}

steps_by_interval <- aggregate(steps ~ interval, data, mean)

plot(steps_by_interval$interval,steps_by_interval$steps, type="l", xlab="5 Min Interval", ylab="Number of Steps",main="Average Number of Steps per Day by Interval (excluding: missing values)")

max_interval <- steps_by_interval[which.max(steps_by_interval$steps),1]
max_interval
```

"835"  5-minute interval, on average across all the days in the dataset, contains the maximum number of steps.



## Imputing missing values

Strategy: Missing values were imputed by inserting the average for each interval.

```{r, echo=TRUE}

incomplete <- sum(!complete.cases(data))
imputed_data <- transform(data, steps = ifelse(is.na(data$steps), steps_by_interval$steps[match(data$interval, steps_by_interval$interval)], data$steps))

steps_by_day_i <- aggregate(steps ~ date, imputed_data, sum)
hist(steps_by_day_i$steps, main = paste("Total Steps Each Day"), col="blue",breaks = 10, xlab="Number of Steps")

NA_Steps <- is.na(data$steps)
 length(NA_Steps)
 
StepMOD_mn <- mean(steps_by_day_i$steps)
StepMOD_md <- median(steps_by_day_i$steps)
StepMOD_mn
StepMOD_md
```

Total number of missing values in the dataset (i.e. the total number of rows with NAs are "17568"

The mean and median total number of steps taken per day are 10766.19 and 10766.19, respectively.

The mean values are identical, but median differs slightly. The modified median is equal to the mean values, and is little higher than the original-non-modified median.

The impact of introducing missing values is not significant for this dataset.



## Are there differences in activity patterns between weekdays and weekends?

```{r, echo=TRUE}

weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")
imputed_data$dow = as.factor(ifelse(is.element(weekdays(as.Date(imputed_data$date)),weekdays), "Weekday", "Weekend"))

steps_by_interval_i <- aggregate(steps ~ interval + dow, imputed_data, mean)

library(lattice)

xyplot(steps_by_interval_i$steps ~ steps_by_interval_i$interval|steps_by_interval_i$dow, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")


```

More steps are taken during weekend. Although, overall, the difference is not much.
